```{r,include=FALSE}
library(plyr)
library(imputeTS)
library(dplyr)
library(tidyverse)
library(forcats)
library(lubridate)
library(aTSA)
library(tseries)
library(forecast)
library(CADFtest)
library(plotly)
library(ggpubr)
library(psych)
library(corrplot)
library(stringr)
library(base)
library(sf)
library(scales)
library(vars)
library(lattice)
library(urca)
library(ggplot2)
library(MASS)
library(haven)
library(joineR)
library(pROC)
library(MESS)
library(arules)
library(fpp2)
```


## Read the data files

The research process involves reading four data files: government stringency_index data, mobility data, weather data, and COVID related data from Our World in data. The goal is to extract valuable information from these files in order to generate a COVID data table. The research aim was to understand the impact of these factors on the spread of COVID-19 and to inform public health decision-making.

```{r}

fulldata <- read.csv('owid_data copy.csv')

weatherNL1=read.csv('weather1.csv')
weatherNL2=read.csv('weather2.csv')

vaccine=read.csv('vaccinations.csv')

weather=rbind(weatherNL1, weatherNL2)
```


```{r}
colnames(fulldata)
```

```{r}
select_date=function(start_date, end_date){

#weather data
weatherNL=weather %>% filter(datetime>=start_date)%>%filter(datetime<= end_date)%>%dplyr::select(datetime, temp, humidity, precip, windspeed)%>%dplyr::rename('date'='datetime')

#aggregating data together
NLdata1 <- fulldata %>%filter(location=="Netherlands") %>% filter(date >= start_date)%>%filter(date<= end_date)%>%dplyr::select(date, new_cases, new_deaths, icu_patients, hosp_patients, stringency_index, reproduction_rate)

df2 <- join_all(list(NLdata1,weatherNL), by='date', type='left')

return(df2)
}
```


```{r}
data_new=select_date('2020-04-01', '2022-11-30')

```

```{r}
data_new$date <- as.Date(data_new$date)
out1=data_new[519:548,c(1, 2) ]
out2=data_new[672:699,c(1, 2) ]
out1$color=ifelse(out1$new_cases==22, 'outlier', 'normal')
out2$color=ifelse(out2$new_cases==2, 'outlier', ifelse(out2$new_cases==380399, 'outlier', 'normal'))
ggplot(out1, )+geom_point(aes(x=date, y=new_cases, color=color))+scale_x_date(date_breaks = "8 day", date_labels =  "%Y-%m-%d")

ggplot(out2, )+geom_point(aes(x=date, y=new_cases, color=color))+scale_x_date(date_breaks = "8 day", date_labels =  "%Y-%m-%d")
```

```{r}
sum(is.na(data_new))
for (i in colnames(data_new)){
  print(i)
  print(sum(is.na(data_new[i])))
}

#[1] "new_deaths"
#[1] 5
#[1] "icu_patients"
#[1] 73
#[1] "hosp_patients"
#[1] 73

```


for the outliers in new cases variable, we replace it using na.interp()
```{r}
data_new$new_cases=ifelse(data_new$new_cases<50, NA,ifelse(data_new$new_cases>300000, NA, data_new$new_cases))
data_new$icu_patients=na.interp(data_new$icu_patients)
data_new$hosp_patients=na.interp(data_new$hosp_patients)
data_new$new_cases=na.interp(data_new$new_cases)
data_new$new_deaths=na.interp(data_new$new_deaths)
data_new$stringency_index=na.interp(data_new$stringency_index)
#data_new$hosp_patients=na.interp(data_new$hosp_patients)
#data_new$people_vaccinated=na.interp(data_new$people_vaccinated)
#data_new$new_vaccinations=na.interp(data_new$new_vaccinations)

data_new$case_fatality=data_new$new_deaths/data_new$new_cases

sum(is.na(data_new))
```

```{r}
describe(data_new)

corrplot(cor(na.omit(data_new[, 2:11])), method = "circle", tl.cex = 0.5) 

```

```{r}

ggplot(data_new, aes(x=date, y=new_cases)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("New cases") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


ggplot(data_new, aes(x=date, y=new_deaths)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("New deaths")+theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")

ggplot( data_new, aes(x=date, y=reproduction_rate)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Reproduction rate") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")

ggplot( data_new, aes(x=date, y=case_fatality)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Case fatality rate") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


ggplot( data_new, aes(x=date, y=hosp_patients)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("hospital admission") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


ggplot(  data_new,aes(x=date, y=icu_patients)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("ICU admission") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


ggplot( data_new, aes(x=date, y=temp)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Temperature") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")

ggplot( data_new, aes(x=date, y=windspeed)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Wind speed") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")
 
ggplot( data_new, aes(x=date, y=humidity)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Humidity") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")

ggplot( data_new, aes(x=date, y=precip)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Rain fall") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


ggplot( data_new, aes(x=date, y=stringency_index)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("Stringency index") +theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))+scale_x_date(date_breaks = "8 month", date_labels =  "%Y-%m")


```





```{r}
data_new$reproduction_rate=data_new$new_deaths/data_new$new_cases
```

```{r}
write.csv(data_new, 'daily_all_data.csv',row.names = FALSE)
```


# aggregating daily data into weekly data
```{r}
week_data=function(df){
   df$week=cut(as.Date(df$date), "week", start.on.monday = TRUE)
   t1=ddply(df, .(week), colwise(mean, 6:11))
   t2=ddply(df, .(week), colwise(sum, 2:5))
   t=join_all(list(t1, t2), by='week', type='left')
  return(t)
}
weekdata=week_data(data_new)
weekdata$case_fatality=weekdata$new_deaths/weekdata$new_cases
```

```{r}
#hosp week t vs week t 
corrplot(cor(na.omit(weekdata[, 2:12])), method = "circle", tl.cex = 1) 

#hosp week t+1 vs week t 
hosp=weekdata$hosp_patients[2:length(weekdata$hosp_patients)]
icu=weekdata$icu_patients[2:length(weekdata$hosp_patients)]
week_lag1=weekdata[1:length(weekdata$date)-1,]
week_lag1$hosp_patients=hosp
week_lag1$icu_patients=icu
#week_lag1=week_lag1[, c(1,12,13,2:11)]
corrplot(cor(na.omit(week_lag1[, 2:12])), method = "circle", tl.cex = 1) 

#hosp week t+2 vs week t 
hosp=weekdata$hosp_patients[3:length(weekdata$hosp_patients)]
icu=weekdata$icu_patients[3:length(weekdata$hosp_patients)]
week_lag1=weekdata[1:length(weekdata$date)-2,]
week_lag1$hosp_patients=hosp
week_lag1$icu_patients=icu
#week_lag1=week_lag1[, c(1,12,13,2:11)]
corrplot(cor(na.omit(week_lag1[, 2:12])), method = "circle", tl.cex = 1) 
```


```{r}
write.csv(weekdata, 'weekly_data.csv',row.names = FALSE)
```



```{r}
sum(is.na(weekdata))
```

